---
title: "Task3 Spark Foundation"
author: "Sannidhya Das"
date: '2023-11-10'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Domain : Data Science and Business Analytics
## Batch : GRIPNOVEMBER23

 Setting the working directory and loading Dataset
```{r}
getwd()
setwd("C:/Users/sanni/OneDrive/Documents/Internship projects/Spark Foundation/Task3")
task3=read.csv("SampleSuperstore.csv")
```

Checking the dataset 
```{r}
head(task3)
tail(task3)
```
Summary of the dataset 
```{r}
summary(task3)
str(task3)
dim(task3)
colnames(task3)
```
Checking is there is any null values in any columns 
```{r}
colSums(is.na(task3))
```

Checking the dataset for duplicates and dropping the duplicate elements using unique() 

```{r}
sum(duplicated(task3))
task3=unique(task3)
```

Finding the correlation and covariance of dataset using cor()and cov() method

```{r}
cor(task3[,c("Sales","Quantity","Discount","Profit")])
cov(task3[,c("Sales","Quantity","Discount","Profit")])

```

 Group the data by multiple columns and calculate the sum of Quantity, Discount, Sales, and Profit

```{r}
grouped <- aggregate(cbind(Quantity, Discount, Sales, Profit) ~ Ship.Mode + Segment + Category + 
                       Sub.Category+ State + Region,
                     data = task3, sum)
```

 Print the grouped data
```{r}
head(grouped)
```
 
 Group the data by State and calculate the sum, mean, min, max, count, median, standard deviation, and variance of Profit

```{r}

library(magrittr)
library(dplyr)
profit_summary <- task3%>% 
  group_by(State) %>% 
  summarise(sum = sum(Profit), mean = mean(Profit), min = min(Profit), max = max(Profit), count = n(),
            median = median(Profit), std = sd(Profit), var = var(Profit))

```
 Print the summary statistics of Profit by State
```{r}
profit_summary
```

## Visualization of dataset

```{r}
library(ggplot2)
ggplot(task3, aes(x = Sub.Category, y = Category)) + 
  geom_bar(stat = "identity", fill = "steelblue") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  ggtitle("Bar plot of Sub-Category vs Category") + 
  xlab("Sub-Category") + 
  ylab("Category")

```
# Note : 
Binders are purchased maximum times from the store followed by papers and phones .

```{r}
library(dplyr)

task3 %>%
  count(State) %>%
  ggplot(aes(x = State, y = n)) +
  geom_bar(stat = "identity", fill = "royalblue") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Count of States") +
  xlab("State") +
  ylab("Count")

```

# Note :
Products are very often ordered from California , New York and Texus .

## Heatmap plot
```{r}
library(reshape2)

# Creating a correlation matrix
corr <- cor(task3[,c("Sales","Quantity","Discount","Profit")])

# Creating a heatmap with annotations
ggplot(melt(corr), aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  geom_text(aes(label = round(value, 2)), color = "black", size = 3) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ggtitle("Correlation Matrix") +
  xlab("Variables") +
  ylab("Variables")

# Creating a covariance matrix
cov <- cov(task3[,c("Sales","Quantity","Discount","Profit")])

# Creating a heatmap with annotations
ggplot(melt(cov), aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  geom_text(aes(label = round(value, 2)), color = "black", size = 3) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ggtitle("Covariance Matrix") +
  xlab("Variables") +
  ylab("Variables")

```

# Note: 
 1.There is a positive Correlation between Sales and   
  profit.(Sales Increase Profit Increases)

 2.There is a positive Correlation between Quantity and  
 Profit.(Quantity Increase Profit Increases)

 3.There is a Negetive Correlation between Profit and
 Discount.(Discount Increase Profit Dicreases)

 4.There is Negative Correlation between Sales and
 Discount.(Sales Increase Discount Decreases)

 5.There is Nearly no Correlation between Quantity and
 Discount.(0 Correlation)



Creating a count plot
```{r}

ggplot(task3, aes(x = Segment, fill = Segment)) +
  geom_bar() +
  scale_fill_manual(values = c("red", "forestgreen", "steelblue")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ggtitle("Count of Segments") +
  xlab("Segment") +
  ylab("Count")
```


```{r}
ggplot(task3, aes(x = Region, fill = Region)) +
  geom_bar() +
  scale_fill_manual(values = c("red", "forestgreen", "steelblue","goldenrod")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ggtitle("Count of Region") +
  xlab("Region") +
  ylab("Count")
```

# Note :
People from Western region orders more products from this store than East , Central and South .


```{r}
ggplot(task3, aes(x = Ship.Mode, fill = Ship.Mode)) +
  geom_bar() +
  scale_fill_manual(values = c("red", "forestgreen", "steelblue","goldenrod")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ggtitle("Count of Ship Mode") +
  xlab("Ship Mode") +
  ylab("Count")
```

# Note : 
When purchasing goods from the store, most customers choose Standard class shipment. 

```{r}
ggplot(task3, aes(x = Category, fill = Category)) +
  geom_bar() +
  scale_fill_manual(values = c("forestgreen", "steelblue","goldenrod")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ggtitle("Count of Category") +
  xlab("Category") +
  ylab("Count")

```

# Note :
People prefers to purchase Office supplies than tech and furniture goods .


# Profit Associated with Sub-Category
```{r}
profit_SubCategory <- aggregate(Profit ~ Sub.Category, data = task3, FUN = sum)

ggplot(profit_SubCategory, aes(x = Sub.Category, y = Profit)) +
  geom_bar(stat = "identity",fill="Maroon") +
  labs(x = "Sub-Category", y = "Profit" ) +
  ggtitle("Profitable Sub-Category") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 
```

#  Note :
 From the above it is clear that Copies gives the maximum profit  to the store . 

# Profit Associated with Segment 
```{r}
profit_Segment <- aggregate(Profit ~ Segment, data = task3, FUN = sum)

ggplot(profit_Segment, aes(x = Segment, y = Profit)) +
  geom_bar(stat = "identity",fill= "brown") +
  facet_wrap(~ Segment, scales = "free_x") +
  labs(x = "Segment", y = "Profit") +
  ggtitle("Profitable Segment") +
  theme(plot.title = element_text(hjust = 0.5))
```

#  Note: 
Consumer give the most profit


# Profit in Association with states 
```{r}
profit_states <- aggregate(Profit ~ State, data = task3, FUN = sum)

ggplot(profit_states, aes(x = State, y = Profit)) +
  geom_bar(stat = "identity",fill="red") +
  labs(x = "State", y = "Profit") +
  ggtitle("Profit in each States")  +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```

#  Note:
California gives maximum profit followed by New York.


Creating a box plot with different colors for different bars
```{r}
ggplot(task3, aes(x = Sub.Category, y = Profit, fill = Sub.Category)) +
  geom_boxplot() +
  scale_fill_manual(values = c("red", "green", "blue", "orange","pink","steelblue","lightgreen","gold","purple","maroon","tan","cadetblue","tomato","orchid","navy","gainsboro","indianred")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ggtitle("Profit by Sub-Category") +
  xlab("Sub-Category") +
  ylab("Profit")
```

Creating a point plot with different colors for different lines
```{r}
ggplot(task3, aes(x = Discount, y = Profit, color = Sub.Category)) +
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ggtitle("Profit by Discount") +
  xlab("Discount") +
  ylab("Profit")
```

Boxplot of Discount
```{r}
ggplot(task3, aes(x = "", y = Discount)) +
  geom_boxplot(color="steelblue") +
  theme(axis.text.x = element_blank(), axis.title.x = element_blank()) +
  ggtitle("Box Plot of Discount") +
  ylab("Discount")
```

Boxplot of profit
```{r}
ggplot(task3, aes(x = "", y = Profit)) +
  geom_boxplot(color="steelblue") +
  theme(axis.text.x = element_blank(), axis.title.x = element_blank()) +
  ggtitle("Box Plot of Profit") +
  ylab("Profit")
```

Distribution Plot
```{r}
summary(task3$Sales)

# Creating a histogram of Sales
hist(task3$Sales, col = "steelblue", breaks = 100, main = "Histogram of Sales", xlab = "Sales", 
     ylab = "Frequency",xlim = c(0,3000))

```

Scatter plot of sales 
```{r}
ggplot(task3, aes(x = Sales, y = Profit)) +
  geom_point(color="seagreen") +
  labs(x = "Sales", y = "Profit") +
  ggtitle("Scatter plot of Sales vs. Profit")

```

Histogram of all quantitative variables 
```{r}
library(tidyr)
h <- gather(task3, key = "variable", value = "value", Sales:Profit)

ggplot(h, aes(x = value)) +
  geom_histogram(bins = 50) +
  facet_wrap(~ variable, scales = "free_x") +
  labs(x = "Value", y = "Frequency") +
  ggtitle("Histogram of Sales, Quantity, Discount, and Profit")


```

# Segment*Region wise profit
```{r}
plot1 <- aggregate(Profit ~ Region + Segment, data = task3, FUN = sum)

ggplot(plot1, aes(x = Region, y = Profit, fill = Segment)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Region", y = "Profit", title = "Comparison of Profit by Region and Segment")  +
  theme(plot.title = element_text(hjust = 0.5))

```

# Note :
Central Region has Less number of consumers

# Profit associated with different categories 
```{r}
plot2 <- aggregate(Profit ~ Region + Category, data = task3, FUN = sum)

ggplot(plot2, aes(x = Region, y = Profit, fill = Category)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Region", y = "Profit", title = "Comparison of Profit by Region and Category") +
  theme(plot.title = element_text(hjust = 0.5))
```

# Note:
1.Sale of furniture is significantly low in Central and Eastern Regions.

2.There is very low office supply in Central Region.

# Lineplots 
# 1.Profit vs Quantity 
```{r}
plot3 <- aggregate(Profit ~ Quantity, data = task3, FUN = sum)

ggplot(plot3, aes(x = Quantity, y = Profit)) +
  geom_line(color = "steelblue") +
  labs(x = "Quantity", y = "Profit", title = "Profit vs. Quantity") +
  theme(plot.title = element_text(hjust = 0.5))

```

# Note:
There is a Constant increment in Profit with the increase in Quantity

# 2. Profit vs Discount 
```{r}
plot4 <- aggregate(Profit ~ Discount, data = task3, FUN = sum)

ggplot(plot4, aes(x = Discount, y = Profit)) +
  geom_line(color = "darkgreen") +
  labs(x = "Discount", y = "Profit", title = "Profit vs. Discount") +
  theme(plot.title = element_text(hjust = 0.5))
```

## Conclusion

1. The superstore loses money when it offers discounts.

2. However, they will lose out on sales and be unable to draw in new, loyal clients if they cease offering discounts.



3. The shop offers discounts around holidays, end-of-season sales, and clearance sales in order to clear up room in their warehouses for new inventory.



4. The business benefits in the long run by gaining more devoted clients by taking on little losses.



5. A crucial aspect of the operation of the corporation is the little losses from discounts.
























